---
title: "How to make plots in R using ggpubr"
author: "Michail Belias"
date: "December 6, 2018"
output:
  slidy_presentation: default
  ioslides_presentation: default
  beamer_presentation: default
---

```{r, echo=FALSE, warning=FALSE, message=FALSE}

if(!require(haven)) install.packages("haven")
if(!require(knitr)) install.packages("knitr")
if(!require(dplyr)) install.packages("dplyr")
if(!require(qwraps2)) install.packages("qwraps2")
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(MASS)) install.packages("MASS")
if(!require(gapminder)) install.packages("gapminder")
if(!require(ggExtra)) install.packages("ggExtra")
if(!require(ggpubr))library(ggpubr)
if(!require(readxl))library(readxl)
if(!require(ggplot2))library(ggplot2)
if(!require(ggsci))library(ggsci)


opts_chunk$set(fig.width=9, fig.height=9, fig.path='Figs/', warning=FALSE, message=FALSE, fig.pos = "H", comment = "")

```


Install and load the packages we will use
========================================================

- You can do it manually through the "install" dialog in the "packages" tab 

```{r echo=FALSE}
if(!require(haven)) install.packages("haven")
if(!require(dplyr)) install.packages("dplyr")
if(!require(ggpubr))library(ggpubr)
if(!require(ggplot2)) install.packages("ggplot2")
if(!require(readxl))library(readxl)
if(!require(gapminder)) install.packages("gapminder")
if(!require(ggExtra)) install.packages("ggExtra")
if(!require(ggsci))library(ggsci)
```



```{r , echo=FALSE, warning=FALSE, message=FALSE}

opts_chunk$set(fig.width=8, fig.height=8, fig.path='Figs/', warning=FALSE, message=FALSE, comment = "")

```

Introduction
========================================================

The ggpubr is a R package that helps you create basic beautiful ggplot2-based graphs.

What is ggpubr:

 * Wrapper around the ggplot2 package for beginners in R programming.
 * Helps researchers, with basic R programming skills, to create easily publication-ready plots.
 * Gives the possibility to add p-values and significance levels to plots.
 * Makes it easy to arrange and annotate multiple plots on the same page.
 * Makes it easy to change grahical parameters such as colors and labels.
 * Is still a ggplot2 object... 
   * Therefore, it can be further manipulated as a ggplot object

First we load the data
========================================================

- We can either use the command line or load them manually
- Use the "Import Dataset" dialog in the Enviroment Tab

```{r}
bladder   <- read_sav("Data/bladder.sav")
surgery   <- read_sav("Data/surgery.sav")
skullrats <- read_sav("Data/SkullRats.sav")

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
surgery$gender =  factor(surgery$gender, labels = c("Male","Female"))
surgery$place =  factor(surgery$place, labels = c(NA,"Outside regional","Local","Regional"))
surgery$infect=  factor(surgery$infect, labels = c("No","Yes"))
surgery$prematur=  factor(surgery$prematur, labels = c("Premature","Term"))
surgery$surgery=  factor(surgery$surgery, labels = c("Abdominal","Cardiac","Other"))


skullrats$treat =  factor(skullrats$treat, labels = c("High dose","Control","Low dose"))


bladder   <- read_sav("Data/bladder.sav")

bladder$gender =  factor(bladder$gender, labels = c("Male","Female"))
bladder$Stage =  factor(bladder$Stage, labels = c("pTa","pT1"))
bladder$Grade=  factor(bladder$Grade, labels = c("Good", "Moderate","Bad"))
bladder$solitaire=  factor(bladder$solitaire, labels = c("Solitaire","Multiple"))
bladder$Died=  as.numeric(bladder$Died)
bladder$Therapy = factor(bladder$Therapy, labels = c("TUR only", "TUR + Chemo","TUR + Radio"))
```

Know your data
========================================================

- Feel free to use your own data in our lab
- with the following commands we can get in touch with our data
  * summary("dataname"), names("dataname"), head("dataname") 
- Use the help tab in your right to see what the functions do (and their options)
- Please play around with the data

Know your data (part2)
========================================================
```{r}
names(bladder)
names(surgery)
names(skullrats)
```


What I want you to learn
========================================================

- Google is our friend
- ggpubr functions have great help file (use it)
- theme function 


Distribution plots
========================================================

Under this section we will show how to plot 

 * Boxplots...........(with ggboxplot) 
 * Histograms.........(with gghistogram)
 * Density Plots......(with ggdensity)

For every function above there is a full description in their help file.
Go to the tab "Help" and write the function in the search box

Boxplot code 
========================================================

First we will see a simple box-plot

```{r}
gg<- ggboxplot(surgery ,  # the data-set
            x = "gender", # the x-values is the categorical variable
            y = "birthwt" # the y-values is their values 
            )
```

Boxplot
========================================================
```{r echo=FALSE}
plot(gg)
```



Full boxplot code 
========================================================

- Search the help file of ggboxplot
- *Tip:* If you select any function and press F1 you go to the function's help file right away

```{r eval=FALSE}
ggboxplot(data, x, y, # load the data and choose the x and y variables
  combine = FALSE, merge = FALSE, color = "black", 
  fill = "white", palette = NULL, title = NULL, xlab = NULL,
  ylab = NULL, bxp.errorbar = FALSE, bxp.errorbar.width = 0.4,
  facet.by = NULL, panel.labs = NULL, short.panel.labs = TRUE,
  linetype = "solid", size = NULL, width = 0.7, notch = FALSE,
  select = NULL, remove = NULL, order = NULL, add = "none",
  add.params = list(), error.plot = "pointrange", label = NULL,
  font.label = list(size = 11, color = "black"), label.select = NULL,
  repel = FALSE, label.rectangle = FALSE, ggtheme = theme_pubr())
```

Full boxplot code 
========================================================

```{r}

  
gg <- ggboxplot(data = surgery,x =  "gender", y = "birthwt",
          xlab = "Gender", ylab = "Birth Weight (in grams)",
          width = 0.5,add = "jitter", 
          shape = "gender",fill = "gender",palette = "simpsons",
          title = "A boxplot of weight at the time of birth",
          legend = "bottom",legend.title="Gender",font.legend = c(10, "bold", "darkgrey"),  
          font.main = c(18, "italic", "black"),
          subtitle = "Male and female infants compared ", font.subtitle = c(12, "bold.italic", "darkgreen"),
          ggtheme = theme_minimal()
) +   
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5)) + 
  stat_compare_means(method = "wilcox.test")#  Add a p-value

```

Boxplot
========================================================
```{r echo=FALSE}
plot(gg)
```

Simple histogram code 
========================================================

```{r}
# Basic histogram plot
gghistogram(surgery, x = "birthwt")
```


Histogram code 
========================================================

```{r}
gg<- gghistogram(surgery, x = "birthwt",y = "..density..",add = "median",
            bins = 10 , # how many bars will the histogram have
            xlab = "Weight at birth (in grams)", ylab = "Density",
            title = "Histogram of infant weight at birth",
            fill = "gender",color = "grey",palette = "lancet",
            alpha = 0.2,legend = "bottom",legend.title="Gender",
            font.legend = c(10, "bold", "darkgrey"), facet.by = "infect", ggtheme = theme_minimal(), add_density = T )+   
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

```

Histogram plot 
========================================================
```{r echo=FALSE}
plot(gg)
```


Simple Density plot code 
========================================================
```{r}
# Basic Density plot 
ggdensity(surgery, x = "birthwt")
```

Density plot code
========================================================

```{r}
gg <-  ggdensity(surgery, x = "birthwt",add = "median",
            xlab = "Weight at birth (in grams)", ylab = "Frequency",
            title = "Density plot of infant weight at birth",
            fill = "gender",color = "grey",palette = "lancet",
            alpha = 0.2,legend = "bottom",legend.title="Gender",
            font.legend = c(10, "bold", "darkgrey"), facet.by = "infect", ggtheme = theme_minimal() )+   
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))
```

Density plot 
========================================================

```{r}
plot(gg)
```




Scatterplots
========================================================

What we learned so far?

1) We use the help file of the gg"function"
2) All the gg functions we:
  - compatible with ggpar() function options
  - produced ggplot2 objects
  - which we can further manipulate using ggplot2 functions
    - theme() , facet_grid() etc are really useful

Scatterplots-Bubble plot code
========================================================
```{r echo=TRUE}
gapminder = gapminder
# Scatterplot
names(gapminder) =  c("Country","Continent","Year","Life_Expectancy",
                      "Population","GDP_per_capita_percentage")

gg = gapminder%>%
    filter(Year %in% "2007")%>%
ggscatter( x = "GDP_per_capita_percentage", y = "Life_Expectancy",
           size = "Population", color = "Continent", 
title = "Life expectancy association with GDP per capita percentage (in 2007)",
           xlab = "GDP per capita" ,ylab = "Life Expectancy",
           legend.position = "right") + 
  theme(plot.title = element_text(hjust = 0.5))
```

Scatterplots Bubble plot 
========================================================

```{r, echo=FALSE}
plot(gg)
```

Scatter-plot code
========================================================

Can we make this plot?
```{r echo=FALSE}
# Scatterplot


gg = ggscatter(data = surgery, x = "birthwt", y = "gestatio", combine = F, merge = T,
          color = "gender",fill = "gender", palette = "lancet", 
          shape = 1, size = 3,point = TRUE, rug = F, 
          title = "Scatterplot of gestational age and birth weight", 
          subtitle = "Faceted by infection information",
          xlab = "Birth weight (in grams)", ylab = "Gestational age",
          facet.by = "infect", 
          panel.labs = list(gender = c("Male", "Female")) , 
          short.panel.labs = TRUE,
          add = "reg.line" , # c("none", "reg.line", "loess")
          conf.int = T, conf.int.level = 0.95, fullrange = T, 
          ggtheme = theme_minimal(),
          label = "id", font.label = c(12, "italic"), font.family = "Comic Sans MS",
          cor.coef = T, cor.coeff.args = list(method = "pearson", ## "pearson", "kendall", or "spearman"
                                              label.x.npc = "left", label.y.npc = "top"),
          cor.coef.coord = c(2000, 42.5), cor.coef.size = 3,show.legend.text = T, 
          repel = T)+   
        theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5))

plot(gg)
# ggMarginal(g, type = "density", fill="transparent")

```


Simple point-line plots
======================================================== 

```{r echo= FALSE}
# Basic plot
ggplot(skullrats,aes(y= response, x = t, group= rat)) + geom_point() + geom_line()
```

Spaggetti plots
======================================================== 

```{r}
ggplot(data=skullrats,aes(x=age,y=response,color=as.factor(rat)))+
  geom_point()+geom_line()+facet_wrap(~treat)+
  theme(legend.title = element_text(size=8),legend.text=element_text(size=7))+
  ylab("Mean Response in pixels")+scale_y_continuous(limits=c(65,92.5),breaks=seq(65,92.5,5),expand=c(0,0))+
  xlab("Age in days")+scale_x_continuous(limits=c(50,110),breaks=seq(50,110,10))+
  scale_color_discrete(name="Rat")+ggtitle("Treatment")+theme(plot.title = element_text(hjust=0.5,size = 10))
```


```{r echo= FALSE}
data1 <- read_excel("Data/Data for graphs.xlsx", 
                    sheet = "graph1")
```


Bar plots code
========================================================

```{r}
gg= ggbarplot(data = data1, #import  Data
          x="App",          # The X-value
          y="Proportion",   # The percentages
          fill="firebrick1",# The color of the bars
          xlab = "",        # label of X-axis
          title = "Bar chart of dating app percentage", # better title
          ylim = c(0,0.80), # Increase the ylimits
          ylab = "Share of respondents",
          order = c("Happn","Lexa","Paiq","Tinder","Badoo"))+   
  theme(plot.title = element_text(hjust = 0.5))+ 
  scale_y_continuous(breaks=seq(0,0.7,0.1),
                     labels=scales::percent)+
  geom_text(aes(label=paste(Proportion*100,"%",sep = "")),size=3,vjust=-1)


```

Bar Chart
========================================================

```{r echo=FALSE}
plot(gg)
```


Ordered bar chart code
========================================================

```{r echo=FALSE}
data2 <- read_excel("Data/Data for graphs.xlsx", 
                    sheet = "Sheet2")

```

```{r}
gg= data2%>%
  arrange(desc(-freq))%>%
  ggbarplot( #import  Data
          x="website",          # The X-value
          y="freq",   # The percentages
          fill="nature",# The color of the bars
          xlab = "",        # label of X-axis
          title = "Online dating services in Netherlands (June 2014)", # better title
          ylim = c(0,300000), # Increase the ylimits
          ylab = "Share of respondents")+ rotate() +
  scale_y_continuous(breaks=seq(0,300000,50000),expand=c(0,0))+
    theme(legend.position =c(0.85,0.25),
          legend.title=element_text(size=8),
          legend.text=element_text(size=8),
          plot.title = element_text(hjust = 0.5))


```

Ordered bar chart 
========================================================
```{r}
plot(gg)
```

Lollipop Chart code
========================================================

```{r}


# Plot
gg= ggplot(data2,aes(x=reorder(website,freq), 
                     y=freq, color=nature, 
                     fill = nature)) + 
  geom_point(size=5,shape = 1) + 
  geom_segment(aes(x=website, xend=website, y=0, yend=freq)) + 
  labs(title="Lollipop Chart", 
       subtitle="Online dating services in Netherlands (June 2014)", 
       caption="source: Alexa.com") + scale_color_lancet()+
  theme(axis.text.x = element_text(angle=65, vjust=0.6))+
  geom_text(aes(label= freq),size=2,hjust=-0.35) + rotate()


```



```{r}
plot(gg)
```

Line-point plots code
========================================================

```{r}
data4 <- read_excel("Data/Data for graphs.xlsx", sheet = "Sheet3")

## Exercise: Now code your plot to look like this:
gg <- ggplot(data=data4,aes(x=year,y=proportion,color=how))+
  geom_point()+geom_line()+theme_minimal()+xlab("Year Couple Met")+ylab("percentage who met this way")+
  scale_y_continuous(limits=c(0,0.7),breaks=seq(0,0.7,0.1),labels=scales::percent,expand=c(0,0))+
  scale_x_continuous(limits=c(1990,2014),breaks=seq(1990,2014,10),expand=c(0,0))+
  theme(legend.position = c(0.2,0.8))+geom_point(aes(shape=how))+scale_shape(guide=FALSE)+
  scale_color_discrete(name="",breaks=c("Met Online","Met through Friends","Bar/Restaurant"))+facet_wrap(~orientation)

```

Line-point plots
========================================================

```{r}
plot(gg)
```




```{r echo=FALSE, message= FALSE}

library(knitr)
purl("presentation.Rpres")

```

